Peer-to-Peer 6502 Cellular Automata

### 2018-07-03 paper.tex
O/S

Each cell has 256 bytes (one page) of volatile state, and a 32K read-only operating system.
All cells' bytes are initially zero, unless otherwise specified
(including, implicitly, cells whose coordinates are outside the CA; see also Section~\ref{sec:AdjacentCells}).
These bytes are the storage of the virtual machine.

@@

There are several emulator-mapped memory addresses that perform special functions:
the write-only addresses \code{SWAP} (\hex{FFF7}) and \code{RULE} (\hex{FFF8}),
and the read-only address \code{RAND} (\hex{FFF9}).

Any value written to \code{SWAP} or \code{RULE} is interpreted as the index of a neighboring cell.
Only the lower 6 bits of the written byte are relevant, and specify a cell in the $3 \times 3 \times 3$ Moore neighborhood
in the same way as bits 8-13 of a memory-mapped storage address (c.f. Table~\ref{tab:AddressBits}).

@@

Writing a byte to \code{RULE} copies the current cell's operating system to the specified neighboring cell
(unless the neighboring cell is outside the CA, in which case it has no effect).

@@

As well as writing the (volatile) state of the cell, you can write its 32K operating system
(mapped to address range \hexrange{8000}{FFEF} in the virtual machine; the last 16 bytes, corresponding to \hexrange{FFF0}{FFFF}, are ignored).
After the first time an operating system is written, it can be referred to by SHA-256 hash, to save space.

@@

% Tables
\begin{table}
\begin{tabular}{llll}
  \hline
  From & To & Type & Contents \\
  \hline
  \hex{0000} & \hex{00FF} & RAM (storage-mapped) & State of cell $(x,y,z)$ \\
  \hex{0100} & \hex{01FF} & RAM (transient) & Stack: initialized to zero \\
  \hex{0200} & \hex{03FF} & RAM (transient) & Scratch memory: initialized to zero \\
  \hex{0400} & \hex{3FFF} & ROM & Reserved: zero \\
  \hex{4000} & \hex{7FFF} & RAM (storage-mapped) & Memory-mapped neighborhood (Table~\ref{tab:MooreNeighborhood}) \\
  \hex{8000} & \hex{FFEF} & ROM & Operating system \\
  \hex{FFF0} & \hex{FFF6} & ROM & Reserved: zero \\
  \hex{FFF7} & \hex{FFF7} & Write-only & \code{SWAP}: swap storage and O/S pages \\
  \hex{FFF8} & \hex{FFF8} & Write-only & \code{RULE}: copy O/S pages to neighbor \\
  \hex{FFF9} & \hex{FFF9} & Read-only & \code{RAND}: random byte generator \\
  \hex{FFFA} & \hex{FFFF} & ROM & Reset and interrupt vectors: zero \\
  \hline
\end{tabular}
\caption{
  \label{tab:MemoryMap}
  Memory map for the 6502 virtual machine of cell $(x,y,z)$.
}
\end{table}

>>>>

Each cell has 256 bytes (one page) of volatile state.
All cells' bytes are initially zero, unless otherwise specified
(including, implicitly, cells whose coordinates are outside the CA; see also Section~\ref{sec:AdjacentCells}).
These bytes are the storage of the virtual machine.

Each cell also has a 32-kilobyte read-only operating system.
Most cells share the same operating system, and operating system changes should be rare,
so the O/S for a cell can be represented using a compact hash key such as SHA-256.

@@

There are several emulator-mapped memory addresses that perform special functions:
the write-only address \code{SWAP} (\hex{FFF8}), which facilitates diffusion models,
and the read-only address \code{RAND} (\hex{FFF9}), which provides a random number generator.

Any value written to \code{SWAP} is interpreted as the index of a neighboring cell.
Only the lower 6 bits of the written byte are relevant, and specify a cell in the $3 \times 3 \times 3$ Moore neighborhood
in the same way as bits 8-13 of a memory-mapped storage address (c.f. Table~\ref{tab:AddressBits}).

@@

Writes to \code{SWAP} are not the only means by which operating system code is propagated across the CA.
During an update, writing any byte to any neighboring cell, i.e. any cell in the memory-mapped storage (\hexrange{4000}{7FFF}),
has the implicit effect of copying the operating system pages from the current cell to the neighbor.
This is handled transparently by the emulator without consuming any additional cycles.

@@

As well as writing the (volatile) state of the cell, you can write its 32K operating system
(mapped to address range \hexrange{8000}{FFF7} in the virtual machine; the last 16 bytes, corresponding to \hexrange{FFF8}{FFFF}, are ignored).
After the first time an operating system is used in the blockchain, it can be referred to by SHA-256 hash, to save space.

@@

% Tables
\begin{table}
\begin{tabular}{llll}
  \hline
  From & To & Type & Contents \\
  \hline
  \hex{0000} & \hex{00FF} & RAM (storage-mapped) & State of cell $(x,y,z)$ \\
  \hex{0100} & \hex{01FF} & RAM (transient) & Stack: initialized to zero \\
  \hex{0200} & \hex{03FF} & RAM (transient) & Scratch memory: initialized to zero \\
  \hex{0400} & \hex{3FFF} & ROM & Reserved: zero \\
  \hex{4000} & \hex{7FFF} & RAM (storage-mapped) & Memory-mapped neighborhood (Table~\ref{tab:MooreNeighborhood}) \\
  \hex{8000} & \hex{FFF7} & ROM & Operating system \\
  \hex{FFF8} & \hex{FFF8} & Write-only & \code{SWAP}: swap storage page with neighbor \\
  \hex{FFF9} & \hex{FFF9} & Read-only & \code{RAND}: random byte generator \\
  \hex{FFFA} & \hex{FFFF} & ROM & Reset and interrupt vectors: zero \\
  \hline
\end{tabular}
\caption{
  \label{tab:MemoryMap}
  Memory map for the 6502 virtual machine of cell $(x,y,z)$.
}
\end{table}
